---
title: "ESS Explore"
author: "Bryan Lockwald"
format: 
  dashboard: 
    orientation: columns
    nav-buttons: [github]
    github: https://github.com/bryanlockwald
sidebar: 
  - title: "Resources"
    items: 
      - text: "Codebook"
        href: documents/ESS_1_9_codebook.html
editor: visual
---

```{r}

library(tidyverse)
library(ggplot2)
library(countrycode)
library(gt)
library(tidyr)
library(DataExplorer)
library(dlookr)







```

```{r}


library(haven)
ess_data <- read_dta("~/Documents/UNLV Political Science PhD/Working Projects/When the Electorate Demands Less Democracy/Data/01_raw/ESS 1-9/ESS1e06_7-ESS2e03_6-ESS3e03_7-ESS4e04_6-ESS5e03_5-ESS6e02_6-ESS7e02_3-ESS8e02_3-ESS9e03_2-ESS10-ESS10SC-ESS11-subset/ESS1e06_7-ESS2e03_6-ESS3e03_7-ESS4e04_6-ESS5e03_5-ESS6e02_6-ESS7e02_3-ESS8e02_3-ESS9e03_2-ESS10-ESS10SC-ESS11-subset.dta")


```

```{r}


## Create new variable of year coresponding to each ESS round

ess_data <- ess_data |> 
  mutate(year = case_when(
    essround == 1 ~ 2002,
    essround == 2 ~ 2004,
    essround == 3 ~ 2006,
    essround == 4 ~ 2008,
    essround == 5 ~ 2010,
    essround == 6 ~ 2012,
    essround == 7 ~ 2014,
    essround == 8 ~ 2016,
    essround == 9 ~ 2018
  )) |> 
  relocate(year, .after = essround)


# create country names from iso2c codes

ess_data <- ess_data  |> 
  mutate(cntry_full = countrycode(sourcevar = cntry, origin = 'iso2c', destination = 'country.name'))

ess_data <- ess_data |> 
  relocate(cntry_full, .after = cntry)


```

# Satisfaction With Institutions

## Row 1

### Satisfaction with Democracy Plot

```{r}


ess_data %>%
  filter(!is.na(stfdem)) %>%
  ggplot(aes(x = year, y = stfdem, fill = cntry)) +
  geom_bar(stat = "summary", fun = "mean", position = "dodge") +
  geom_smooth(method = "lm", color = "red", se = FALSE, aes(group = cntry)) +
  labs(title = "Satisfaction with Democracy Over ESS Rounds",
       subtitle = "ESS rounds are published every two years",
       x = "ESS Round",
       y = "Mean Satisfaction with Democracy") +
  facet_wrap(~cntry_full) +
  theme_minimal() +
  theme(legend.position = "none")


```

### Satisfaction with Government

```{r}

ggplot(ess_data, aes(x = year, y = stfgov, fill = cntry_full)) +
  geom_bar(stat = "summary", fun = "mean", position = "dodge") +
  geom_smooth(method = "lm", color = "red", se = FALSE, aes(group=cntry)) +
  labs(title = "Satsifaction with Government - ESS Rounds Years 2002 - 2018",
       x = "ESS Round",
       y = "Mean Satisfaction with Government") +
  facet_wrap(~cntry_full) +
  theme_minimal() +
  theme(legend.position = "none")




```

### Satisfaction with Economy

```{r}

ggplot(ess_data, aes(x = year, y = stfeco, fill = cntry_full)) +
  geom_bar(stat = "summary", fun = "mean", position = "dodge") +
  geom_smooth(method = "lm", color = "red", se = FALSE, aes(group=cntry)) +
  labs(title = "Satsifaction with Economy - ESS Rounds Years 2002 - 2018",
       x = "ESS Round",
       y = "Mean Satisfaction with Economy") +
  facet_wrap(~cntry_full) +
  theme_minimal() +
  theme(legend.position = "none")

```

## Row 2

### Satisfaction with Education

```{r}

ggplot(ess_data, aes(x = year, y = stfedu, fill = cntry_full)) +
  geom_bar(stat = "summary", fun = "mean", position = "dodge") +
  geom_smooth(method = "lm", color = "red", se = FALSE, aes(group=cntry)) +
  labs(title = "Satsifaction with Education - ESS Rounds Years 2002 - 2018",
       x = "ESS Round",
       y = "Mean Satisfaction with Education") +
  facet_wrap(~cntry_full) +
  theme_minimal() +
  theme(legend.position = "none")

```

### State of Health Services

```{r}

ggplot(ess_data, aes(x = year, y = stfhlth, fill = cntry_full)) +
  geom_bar(stat = "summary", fun = "mean", position = "dodge") +
  geom_smooth(method = "lm", color = "red", se = FALSE, aes(group=cntry)) +
  labs(title = "Satsifaction with Health Services - ESS Rounds Years 2002 - 2018",
       x = "ESS Round",
       y = "Mean Satisfaction with Health Services") +
  facet_wrap(~cntry_full) +
  theme_minimal() +
  theme(legend.position = "none")

```

### Life Satisfaction

```{r}

ggplot(ess_data, aes(x = year, y = stflife, fill = cntry_full)) +
  geom_bar(stat = "summary", fun = "mean", position = "dodge") +
  geom_smooth(method = "lm", color = "red", se = FALSE, aes(group=cntry)) +
  labs(title = "Satsifaction with Life - ESS Rounds Years 2002 - 2018",
       x = "ESS Round",
       y = "Mean Satisfaction with Life") +
  facet_wrap(~cntry_full) +
  theme_minimal() +
  theme(legend.position = "none")

```

# Trust in Institutions

## Row 1 Institutions

### Trust in European Parliament

```{r}

ggplot(ess_data, aes(x = year, y = trstep, fill = cntry_full)) +
  geom_bar(stat = "summary", fun = "mean", position = "dodge") +
  geom_smooth(method = "lm", color = "red", se = FALSE, aes(group=cntry)) +
  labs(title = "Trust in EU Perliament - ESS Rounds Years 2002 - 2018",
       x = "ESS Round",
       y = "Mean Trust in EU Parliament") +
  facet_wrap(~cntry_full) +
  theme_minimal() +
  theme(legend.position = "none")

```

### Trust in Legal System

```{r}

ggplot(ess_data, aes(x = year, y = trstlgl, fill = cntry_full)) +
  geom_bar(stat = "summary", fun = "mean", position = "dodge") +
  geom_smooth(method = "lm", color = "red", se = FALSE, aes(group=cntry)) +
  labs(title = "Trust in the Legal System - ESS Rounds Years 2002 - 2018",
       x = "ESS Round",
       y = "Mean Trust in Legal System") +
  facet_wrap(~cntry_full) +
  theme_minimal() +
  theme(legend.position = "none")

```

### Trust in Police

```{r}

ggplot(ess_data, aes(x = year, y = trstplc, fill = cntry_full)) +
  geom_bar(stat = "summary", fun = "mean", position = "dodge") +
  geom_smooth(method = "lm", color = "red", se = FALSE, aes(group=cntry)) +
  labs(title = "Trust in the Police - ESS Rounds Years 2002 - 2018",
       x = "ESS Round",
       y = "Mean Trust in Police") +
  facet_wrap(~cntry_full) +
  theme_minimal() +
  theme(legend.position = "none")
```

## Row 2 Institutioins

### Trust in Polititians

```{r}

ggplot(ess_data, aes(x = year, y = trstplt, fill = cntry_full)) +
  geom_bar(stat = "summary", fun = "mean", position = "dodge") +
  geom_smooth(method = "lm", color = "red", se = FALSE, aes(group=cntry)) +
  labs(title = "Trust in Politicians - ESS Rounds Years 2002 - 2018",
       x = "ESS Round",
       y = "Mean Trust in Politicians") +
  facet_wrap(~cntry_full) +
  theme_minimal() +
  theme(legend.position = "none")

```

### Trust in Parliament

```{r}

ggplot(ess_data, aes(x = year, y = trstprl, fill = cntry_full)) +
  geom_bar(stat = "summary", fun = "mean", position = "dodge") +
  geom_smooth(method = "lm", color = "red", se = FALSE, aes(group=cntry)) +
  labs(title = "Trust in National Parliament - ESS Rounds Years 2002 - 2018",
       x = "ESS Round",
       y = "Mean Trust in National Parliament") +
  facet_wrap(~cntry_full) +
  theme_minimal() +
  theme(legend.position = "none")

# ess_data %>%
#   filter(!is.na(trstprl)) %>%
#   ggplot(aes(x = year, y = trstprl, fill = cntry_full)) +
#   geom_bar(stat = "summary", fun = "mean", position = "dodge") +
#   geom_smooth(method = "lm", color = "red", se = FALSE, aes(group = cntry)) +
#   labs(title = "Trust in National Parliament - ESS Rounds Years 2002 - 2018",
#        x = "ESS Round",
#        y = "Mean Trust in National Parliament") +
#   facet_wrap(~cntry_full) +
#   theme_minimal() +
#   theme(legend.position = "none")

```

### Trust in the United Nations

```{r}

ggplot(ess_data, aes(x = year, y = trstun, fill = cntry_full)) +
  geom_bar(stat = "summary", fun = "mean", position = "dodge") +
  geom_smooth(method = "lm", color = "red", se = FALSE, aes(group=cntry)) +
  labs(title = "Trust in The United Nations - ESS Rounds Years 2002 - 2018",
       x = "ESS Round",
       y = "Mean Trust in the UN") +
  facet_wrap(~cntry_full) +
  theme_minimal() +
  theme(legend.position = "none")




```

# Immigration

## Row 1 Immigration

### Allow few/many of same race/ethnic group as majority?

```{r}

ggplot(ess_data, aes(x = year, y = imsmetn, fill = cntry_full)) +
  geom_bar(stat = "summary", fun = "mean", position = "dodge") +
  geom_smooth(method = "lm", color = "red", se = FALSE, aes(group=cntry)) +
  labs(title = "Allow few/many immigrants of same race/ethnic group as majority?",
       x = "ESS Round",
       y = "Mean Allow same Race/Ethnic Groups?") +
  facet_wrap(~cntry_full) +
  theme_minimal() +
  theme(legend.position = "none")

```

### Allow few/many of different race/ethnic group as majority?

```{r}

ggplot(ess_data, aes(x = year, y = imdfetn, fill = cntry_full)) +
  geom_bar(stat = "summary", fun = "mean", position = "dodge") +
  geom_smooth(method = "lm", color = "red", se = FALSE, aes(group=cntry)) +
  labs(title = "Allow few/many immigrants of different race/ethnic group as majority?",
       x = "ESS Round",
       y = "Mean Allow different Race/Ethnic Groups?") +
  facet_wrap(~cntry_full) +
  theme_minimal() +
  theme(legend.position = "none")

```

### Allow many/few immigrants from poorer countries outside of Europe?

```{r}

ggplot(ess_data, aes(x = year, y = impcntr, fill = cntry_full)) +
  geom_bar(stat = "summary", fun = "mean", position = "dodge") +
  geom_smooth(method = "lm", color = "red", se = FALSE, aes(group=cntry)) +
  labs(title = "Allow many/few immigrants from poorer countries outside of Europe?",
       x = "ESS Round",
       y = "Mean Allow Immigrants from Poor Countries?") +
  facet_wrap(~cntry_full) +
  theme_minimal() +
  theme(legend.position = "none")
```

## Row 2 Immigration

### Immigration good/bad for countries economy?

```{r}

ggplot(ess_data, aes(x = year, y = imbgeco, fill = cntry_full)) +
  geom_bar(stat = "summary", fun = "mean", position = "dodge") +
  geom_smooth(method = "lm", color = "red", se = FALSE, aes(group=cntry)) +
  labs(title = "Immigration good/bad for countries economy?",
       x = "ESS Round",
       y = "Immigration good/bad for countries economy?") +
  facet_wrap(~cntry_full) +
  theme_minimal() +
  theme(legend.position = "none")

```

### Countries Cultural Life is Undermined or Enriched by Immigrants?

```{r}

ggplot(ess_data, aes(x = year, y = imueclt, fill = cntry_full)) +
  geom_bar(stat = "summary", fun = "mean", position = "dodge") +
  geom_smooth(method = "lm", color = "red", se = FALSE, aes(group=cntry)) +
  labs(title = "Countries Cultural Life is Undermined or Enriched by Immigrants?",
       x = "ESS Round",
       y = "Undermined or Enriched by Immigrants?") +
  facet_wrap(~cntry_full) +
  theme_minimal() +
  theme(legend.position = "none")

```

# Code Book

[Codebook](https://bryanlockwald.github.io/DemosParadox/)
